---
title: "CaseStudy2DDS"
author: "Shiju Raju"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## R Markdown

This R markdown file contains codes to analyze the "CaseStudy2.data" dataset
to identify factors that lead to Attrition.

36 variables that are given in the data set were analyzed to identify the 
top three factors that contribute to turnover(backed up by evidence provided
by the analysis). Otherjob role specific trends were also analysed for this 
project. Prediction of Attrition and Salary were also conducted using
different machine learning techniques such as Nearest Neigbor, Naive Bayes and
Linear Regression.


```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(GGally)
#library(GGpairs)

CaseStudy2.data<- read.csv("C:/Users/sheena/Desktop/SMU_Data_Science/Week14-15_Project2/CaseStudy2-data_try1.csv")

#Find out how many out of 870 is Attrition.
CaseStudy2.data%>%ggplot(aes(x=Attrition,fill= Attrition))+
  geom_bar(position="dodge")+ggtitle("Attrition")+
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white")
#Filter out people stayed and left
left=CaseStudy2.data%>%filter(Attrition=="Yes")
stayed=CaseStudy2.data%>%filter(Attrition=="No")
```

```{r}
#Age-Mean, Median of people who left
summary(left$Age)

#Age-Mean, Median of people who stayed
summary(stayed$Age)

#Distribution of Age of people who left
left %>% select(Age, Attrition) %>%  ggpairs(aes(color = Attrition))
```

```{r}
#--------------------------------2. Job Satisfaction----------------------------
CaseStudy2.data %>%
  count(JobSatisfaction, Attrition) %>%       
  group_by(JobSatisfaction) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(JobSatisfaction, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Job Satisfaction & Attrition") +
  theme_bw()
```

```{r}
#------------------------------3. Relationship Satisfaction---------------------
# Relationship satisfaction of the people who left
CaseStudy2.data %>%
  count(RelationshipSatisfaction, Attrition) %>%       
  group_by(RelationshipSatisfaction) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(RelationshipSatisfaction, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Relationship Satisfaction & Attrition") +
  theme_bw()
```

```{r}
#------------------------------4. Monthly Income--------------------------------
Monthly_Income= cut(CaseStudy2.data$MonthlyIncome, breaks = c(0,5000,10000,15000,20000), labels = c("Low","Medium","High","Very High"))
  CaseStudy2.data%>% mutate(Monthly_Income = Monthly_Income)%>% ggplot(aes(x = MonthlyIncome,  y=Attrition,  color = Monthly_Income)) +
   geom_point() + ggtitle("Attrition vs Monthly Income level")+
    xlab("Monthly Income ($)") + ylab("Attrition")

summary(left$MonthlyIncome)
summary(stayed$MonthlyIncome)
t.test(CaseStudy2.data$MonthlyIncome~CaseStudy2.data$Attrition)
#---------------------------------5.Education Level-----------------------------
CaseStudy2.data %>%
  count(Education, Attrition) %>%       
  group_by(Education) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(Education, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Education level & Attrition") +
  theme_bw()


#------------------------6. Commuting Distance----------------------------------
CaseStudy2.data %>%
  count(DistanceFromHome, Attrition) %>%       
  group_by(DistanceFromHome) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(DistanceFromHome, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Commuting Distance & Attrition") +
  theme_bw()

#------------------------7. Gender----------------------------------------------
CaseStudy2.data %>%
  count(Gender, Attrition) %>%       
  group_by(Gender) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(Gender, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Gender & Attrition") +
  theme_bw()
#---------------------------------8.Education Field-----------------------------
CaseStudy2.data %>%
  count(EducationField, Attrition) %>%       
  group_by(EducationField) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(EducationField, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Education Field & Attrition") +
  theme_bw()

#---------------------------------9.Marital Status------------------------------
CaseStudy2.data %>%
  count(MaritalStatus, Attrition) %>%       
  group_by(MaritalStatus) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(MaritalStatus, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("MaritalStatus & Attrition") +
  theme_bw()

#---------------------------------10.Work Life Balance--------------------------
CaseStudy2.data %>%
  count(WorkLifeBalance, Attrition) %>%       
  group_by(WorkLifeBalance) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(WorkLifeBalance, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Work-Life Balance & Attrition") +
  theme_bw()

#-----------------11.Years Since Last Promotion---------------------------------
CaseStudy2.data %>%
  count(YearsSinceLastPromotion, Attrition) %>%       
  group_by(YearsSinceLastPromotion) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(YearsSinceLastPromotion, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Years Since Last Promotion & Attrition") +
  theme_bw()

#-----------------12.Years in Current Role--------------------------------------
CaseStudy2.data %>%
  count(YearsInCurrentRole, Attrition) %>%       
  group_by(YearsInCurrentRole) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(YearsInCurrentRole, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Years In Current Role & Attrition") +
  theme_bw()

#------------------------------13.Over Time-------------------------------------
CaseStudy2.data %>%
  count(OverTime, Attrition) %>%       
  group_by(OverTime) %>%
  mutate(pct= prop.table(n) * 100) %>%
  ggplot() + aes(OverTime, pct, fill=Attrition) +
  geom_bar(stat="identity") +
  ylab("Percentage") +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct),"%")),
            position=position_stack(vjust=0.5)) +
  ggtitle("Over Time & Attrition") +
  theme_bw()
#------------------------14.Education Level & Income----------------------------

CaseStudy2.data%>% mutate(Monthly_Income = Monthly_Income)%>% ggplot(aes(x = MonthlyIncome,  y= Education,  color= Monthly_Income)) +
  geom_point() + ggtitle("Education vs Monthly Income level")+ 
  xlab("Monthly Income $") + ylab("Education")

```


```{r}
#-------------------------------------------------------------------------------
#------------------------Part 2. Job Specific Trends----------------------------
#-------------------------------------------------------------------------------
#Job Role and JobSatisfaction
ggplot(data = CaseStudy2.data, mapping = aes(x = EducationField,y=JobSatisfaction)) +
  geom_point() +ggtitle("Education Field and JobSatisfaction")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Job Role and Monthly Income
ggplot(data = CaseStudy2.data, mapping = aes(x = JobRole,y=MonthlyIncome)) +
  geom_point() +ggtitle("Job Role and Monthly Income")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Job Role and Years at a Company
ggplot(data = CaseStudy2.data, mapping = aes(x = JobRole,y=YearsAtCompany)) +
  geom_point() +ggtitle("Job Role and Years at a Company")+xlab("JobRole")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Age and Monthly Income
ggplot(data = CaseStudy2.data, mapping = aes(x = Age,y=Monthly_Income)) +
  geom_point() +ggtitle("Age and Monthly Income")

#-------------------------------------------------------------------------------
#------------------------Model to predict Attrition-KNN-------------------------
#-------------------------------------------------------------------------------
library(class)
library(caret)
library(e1071)

set.seed(6)
splitPerc = .70

trainIndices = sample(1:dim(CaseStudy2.data)[1],round(splitPerc * dim(CaseStudy2.data)[1]))
train = CaseStudy2.data[trainIndices,]
test = CaseStudy2.data[-trainIndices,]

CaseStudy2.data %>% ggplot(aes(x = MonthlyIncome,JobSatisfaction,color = Attrition)) + geom_point()+
  ggtitle("Attrition based on Monthly Income & Job Satisfaction") + 
  xlab("Monthly Income($)") + ylab("Job Satisfaction")

## Loop for many k and one training / test partition
accs = data.frame(accuracy = numeric(90), k = numeric(90))
for(i in 1:90)
{
  classifications = knn(train[,c(18,20)],test[,c(18,20)],train$Attrition, prob = TRUE, k = i)
  table(test$Attrition,classifications)
  CM = confusionMatrix(table(test$Attrition,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
plot(accs$k,accs$accuracy, type = "l", xlab = "k")
CM

df_knn=data.frame(table(test$ID,test$Attrition))
#changing variable name
names(df_knn)[1]<-"ID"
names(df_knn)[2]<-"Attrition"
names(df_knn)[3]<-"Count"
write.csv(df_knn, file = "C:\\Users\\sheena\\Desktop\\SMU_Data_Science\\Week14-15_Project2\\Case2PredictionsRAJU_Attrition", row.names = FALSE)

#---------------------------------WLB & MonthlyINcome---------------------------
set.seed(6)
splitPerc = .70

trainIndices = sample(1:dim(CaseStudy2.data)[1],round(splitPerc * dim(CaseStudy2.data)[1]))
train = CaseStudy2.data[trainIndices,]
test = CaseStudy2.data[-trainIndices,]

CaseStudy2.data %>% ggplot(aes(x = MonthlyIncome,WorkLifeBalance,color = Attrition)) + geom_point()+
  ggtitle("Attrition based on Monthly Income & Work Life Balance") + 
  xlab("Monthly Income($)") + ylab("Work Life Balance")

## Loop for many k and one training / test partition
accs = data.frame(accuracy = numeric(90), k = numeric(90))
for(i in 1:90)
{
  classifications = knn(train[,c(32,20)],test[,c(32,20)],train$Attrition, prob = TRUE, k = i)
  table(test$Attrition,classifications)
  CM = confusionMatrix(table(test$Attrition,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
plot(accs$k,accs$accuracy, type = "l", xlab = "k")
CM

df=data.frame(table(test$ID,test$Attrition))
#changing variable name
names(df)[1]<-"ID"
names(df)[2]<-"Attrition"
names(df)[3]<-"Count"

#----------------Model to predict Attrition-Naive Bayes-------------------------
set.seed(7)
trainIndices = sample(seq(1:length(CaseStudy2.data$MonthlyIncome)),round(.7*length(CaseStudy2.data$JobSatisfaction)))
trainAttriton = CaseStudy2.data[trainIndices,]
testAttrition = CaseStudy2.data[-trainIndices,]

model_train = naiveBayes(Attrition~.,data = trainAttriton)
predict(model_train, trainAttriton[,c(18,20)],)

model_test = naiveBayes(Attrition~.,data = testAttrition)
classifications=predict(model_test, testAttrition[,c(18,20)])
CM_NB=confusionMatrix(table(classifications,testAttrition$Attrition))
CM_NB

df2=data.frame(table(test$ID,test$Attrition))
#changing variable name
names(df2)[1]<-"ID"
names(df2)[2]<-"Attrition"
names(df2)[3]<-"Count"


#-------------------------------------------------------------------------------
#------------------------Model to predict Salary-Regression---------------------
#-------------------------------------------------------------------------------
#Age & Monthly_Income
ggplot(data = CaseStudy2.data, mapping = aes(x = Age,y=MonthlyIncome)) +
  geom_point() +ggtitle("Age and Monthly Income")

#degree 1 model
fit = lm(Age~MonthlyIncome, data = CaseStudy2.data)
summary(fit)
fit$coefficients
fit$coefficients[1] 
fit$coefficients[2]

preds = predict(fit)

CaseStudy2.data %>%  ggplot(aes(x = Age, y = MonthlyIncome)) + geom_point() +
  ggtitle("Age vs. Monthly Income") + geom_smooth(method = "lm")

df3=data.frame(table(test$ID,test$MonthlyIncome))
#changing variable name
names(df3)[1]<-"ID"
names(df3)[2]<-"Attrition"
names(df3)[3]<-"Count"

write.csv(df3, file = "C:\\Users\\sheena\\Desktop\\SMU_Data_Science\\Week14-15_Project2\\Case2PredictionsRAJU_Salary", row.names = FALSE)

#degree 2 model
CaseStudy2.data %>% ggplot(aes(x = Age, y = MonthlyIncome)) + geom_point()
df4 = CaseStudy2.data %>% mutate(Age2 = Age^2)
fit = lm(MonthlyIncome~Age+Age2, df4)
summary(fit)

preds = predict(fit)
CaseStudy2.data %>% ggplot(aes(x = Age, y = MonthlyIncome)) + geom_point() +
  geom_line(data = CaseStudy2.data, aes( x = Age, y = preds, col = "red"))

#degree 3 model
CaseStudy2.data %>% ggplot(aes(x = Age, y = MonthlyIncome)) + geom_point()
df5 = CaseStudy2.data %>% mutate(Age2=Age^2, Age3 = Age^3)
fit = lm(MonthlyIncome~Age+Age2+Age3, df5)
summary(fit)

preds = predict(fit)
CaseStudy2.data %>% ggplot(aes(x = Age, y = MonthlyIncome)) + geom_point() +
  geom_line(data = CaseStudy2.data, aes( x = Age, y = preds, col = "red"))

#Youtube Link
#https://www.youtube.com/watch?v=RW-12d72Ino

#Techsmith Link
#https://www.screencast.com/t/am9Jtjn7
#------------------------------------END PROJECT-------------------------------

```
